---
title: "FinalProject"
output: html_document
date: "2022-11-22"
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
data<-read.csv("Maths.csv")
library(lattice)
library(reticulate)
library(tidyverse)
library(dplyr)
library(MLmetrics)
library(MASS)
```

# Data Preparation

## Summary
```{r}
summary(data)
```

## Dimensions
```{r}

dim(data)

```

## Missing Values
```{r}

sum(is.na(data))
```

## Type of data
```{r}

print(str(data))

```



# Data Exploration(EDA)

## Histogram 

We will do some histograms to have a better understanding of the data used. 

```{r}

par(mfrow = c(2, 2))

#Histogram of Petal Width with 10 bins
hist(data$G1,
          main="Grade First Semerter",
          xlab="Grade")

hist(data$G2,
          main="Grade Second Semester",
          xlab="Grade")

hist(data$G3,
          main="Overall grade",
          xlab="Grade")

hist(data$age,
          main="Age",
          xlab="Age")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Barplot

Since we have a lot of columns with defined variables, it is interesting to make some barplots so we can get a better uderstanding of the dataset.

### Barplots related to studying

```{r}

par(mfrow = c(2, 2))

barplot(table(data$higher),  main="Students who wants to get a higher education")
barplot(table(data$studytime), main = "Amount of study time")
barplot(table(data$activities), main = "Extra-curricular activities")
barplot(table(data$paid), main = "Paid courses")

```
> Comments 

### Barplots related to drinking

```{r}

par(mfrow = c(2, 2))
barplot(table(data$freetime),  main="Amount of free time")
barplot(table(data$goout), main = "Going out with friends")
barplot(table(data$Dalc), main = "Workday alcohol consumption")
barplot(table(data$Walc), main = "Weekend alcohol consumption")

```
> Comment

## Boxplots 

Now that we undertand our dataset a little bit, we will do some boxplots to analyze is some of our data have outliers and/or if the values are spread out or not.


```{r}
par(mfrow = c(1, 2))

boxplot(data$absences, main = "Absences")
summary(data$absences)
boxplot(data$age, main = "Age")
summary(data$age)
```

## Histogram with overlay

```{r}

ggplot(data = data, aes(x = Walc)) +
  geom_histogram(aes(fill = sex), bins = 30, color = "black") 


```

## Other graphs with overlay

```{r}

ggplot(data, aes(x = studytime, y = Walc)) +
  geom_count(alpha = 0.5)

ggplot(data, aes(x = G3, y = Walc)) +
  geom_count(alpha = 0.5)
```

# Data Analysis

## Hypothesis Testing
```{r}
alcohol <- data[c('age','failures', 'sex', 'G2', 'G3', 'Dalc', 'Walc', 'studytime', 'failures', 'goout')]
str(alcohol)
```

### Effect of alcohol on grades

h0: Mean of grade of people who drink more = Mean of grade of people who drink less
h1: Mean of people who drink more != Mean of people who drink less

```{r}
less <- alcohol[alcohol['Walc'] < 3,]
more <- alcohol[alcohol['Walc'] > 2,]
var.test(more$G3, less$G3)
t.test(more$G3, less$G3)
```

### Effect of study time on grades

h0: Mean of grades of people who study more = Mean of grades of people who study less
h1: Mean of grades of people who study more != Mean of grades of people who study less

```{r}
less <- alcohol[alcohol['studytime'] < 3,]
more <- alcohol[alcohol['studytime'] > 2,]
var.test(more$G3, less$G3)
t.test(more$G3, less$G3, var.equal = FALSE)
```
### Relation between failure and drinking

h0: Mean of alcohol of people who fail more = Mean of alcohol of people who fail less
h1: Mean of alcohol of people who fail more != Mean of alcohol of people who fail less

```{r}
less <- data[data['failures'] < 3,]
more <- data[data['failures'] > 2,]
t.test(more$Walc, less$Walc)

less <- data[data['absences'] < 46,]
more <- data[data['absences'] > 47,]
t.test(more$Walc, less$Walc)
```
### Relation between gender and drinking

h0: Mean of alcohol of woman = Mean of alcohol of men
h1: Mean of alcohol of woman != Mean of alcohol of men

```{r}
female <- alcohol[alcohol['sex'] == 'F',]
male <- alcohol[alcohol['sex'] == 'M',]
t.test(female$Walc, male$Walc)
```

## Linear regression model

### Forward Stepwise

```{r}
# Create a null model 
intercept_only <- lm(G3 ~ 1, data=data)
# Create a full model
all <- lm(G3~., data=data)
# perform forward step-wise regression
forward <- stepAIC (intercept_only, direction='forward',scope = formula(all))
forward$anova
```

### Model
```{r}
summary(forward)
```
```{r}
plot(forward)
train.control <- trainControl(method = "cv", number = 10)
forward.model <- train(Fertility ~., data = data,
                    method = "lmStepAIC", 
                    trControl = train.control,
                    trace = FALSE
                    )
# Model accuracy
```

### MAE and MSE
```{r}
dataM<-read.csv("Portuguese.csv")
dataM.predict <- predict(object = forward, newdata = dataM)

diff <- abs(dataM$G3 - dataM.predict)

print(mean(diff))

MAE(y_pred = ypred_forward, y_true = data$G3)
plot(ypred_forward)
```
```{r}
MSE(y_pred = ypred_forward, y_true = data$G3)
```

